---
title: "Data"

format:
  html:
    code-fold: true
    code-summary: "Show the code"
---

Datasets shown below downloaded from [Istanbul Metropolitan Municipality](data.ibb.gov.tr) website as csv and xlsx files. Separate links are also provided.

# Data Sets

1.  [Earthquake Scenario Analysis Results](https://data.ibb.gov.tr/tr/dataset/deprem-senaryosu-analiz-sonuclari/resource/9c3ac492-de4b-4245-b418-7ad3df67a193)

    In this dataset some column names has been changed. Structure and first 10 observation can be seen below. Download [R.Data](https://github.com/emu-hacettepe-analytics/emu430-fall2023-team-team_safe_istanbul/blob/master/deprem_analiz.RData) file to see the entire dataframe.

    ```{r, message=FALSE}
    library(dplyr)
    library(readr)
    library(knitr)
    deprem_analiz <- read.csv("Deprem_senaryosu_analiz_sonuçlar.csv")
    colnames(deprem_analiz)[1:4] <- c("id", "ilce_adi", "mahalle_adi", "mahalle_kodu")
    kable(str(deprem_analiz), caption="Structure of deprem_analiz")
    save(deprem_analiz, file = "deprem_analiz.RData")
    ```

2.  [Neighborhood-Based Building Numbers in 2017](https://data.ibb.gov.tr/tr/dataset/mahalle-bazli-bina-analiz-verisi/resource/cef193d5-0bd2-4e8d-8a69-275c50288875)

    **toplam_bina** column is calculated and added to the last. Also, some colum names has been changed.

    ```{r, message=FALSE}
    mahalle_bazli_bina <- read.csv("mahalle_bazli_bina_2017.csv")
    colnames(mahalle_bazli_bina) <- c("id", "ilce_adi", "mahalle_adi", "mahalle_kodu", "once_1980", "ara_1980_2000", "sonra_2000", "ara_1_4_kat", "ara_5_9_kat", "ara_9_19_kat")
    toplam_bina <- mahalle_bazli_bina %>% 
      select(ara_1_4_kat,ara_5_9_kat,ara_9_19_kat) %>%
      rowwise() %>%
      do( (.) %>% as.data.frame() %>% mutate(toplam_bina = sum(.))) %>%
      ungroup %>%
      select(toplam_bina)
    mahalle_bazli_bina <- cbind(mahalle_bazli_bina, toplam_bina)
    str(mahalle_bazli_bina)
    kable(head(mahalle_bazli_bina, n = 10L), caption="Head of Mahalle Bazlı Bina")
    ```

    A new column has been created to examine buildings in districts separately. You can see the first 10 district that have highest number of buildings.

    ```{r}
    kable(mahalle_bazli_bina %>% 
      group_by(ilce_adi) %>%
      summarize(ilce_bazli_bina = sum(toplam_bina)) %>%
      arrange(desc(ilce_bazli_bina)) %>%
      head(n = 10L))
    ```

    ```{r}
    save(mahalle_bazli_bina, file = "mahalle_bazli_bina.RData")
    ```

    Refer to [R.Data](https://github.com/emu-hacettepe-analytics/emu430-fall2023-team-team_safe_istanbul/blob/master/mahalle_bazli_bina.RData) file.

3.  [Municipality Population in 2019](https://data.ibb.gov.tr/tr/dataset/belediye-nufuslari-veri-seti/resource/c6c9b289-2824-41b3-ab3d-4fd655ed4e24)

    Structure and first 10 observation can be seen below.

    ```{r}

    library(readxl)
    nufus <- read_excel("belediye_nufuslar_2019.xlsx")
    str(nufus)
    head(nufus, n = 10L)
    save(nufus, file = "nufus.RData")
    ```

    You can also download this dataset as [R.Data](https://github.com/emu-hacettepe-analytics/emu430-fall2023-team-team_safe_istanbul/blob/master/nufus.RData) file.

    **An Overview**

    ```{r}
    library(ggplot2)
    vis <- deprem_analiz %>% 
      inner_join(mahalle_bazli_bina, by = "id") %>%
      select(ilce_adi.x, ara_1_4_kat, ara_5_9_kat, ara_9_19_kat, toplam_bina, can_kaybi_sayisi) 
       
    vis %>% ggplot(aes(x = toplam_bina, y = can_kaybi_sayisi, color = ilce_adi.x )) + geom_abline(slope = log10(10)/log10(100),intercept = 0) +
      scale_x_log10()+
      scale_y_log10()+
      geom_point()
      
    ```

    **The plot shown above shows us total loss vs total number of buildings.**

    ```{r}

    vis2 <- deprem_analiz %>% 
      inner_join(mahalle_bazli_bina, by = "id") %>%
      select(ilce_adi.x,once_1980, ara_1980_2000, sonra_2000, toplam_bina, can_kaybi_sayisi,toplam_bina) %>% 
      group_by(ilce_adi.x) %>%
      summarize(toplam_once_1980 = sum(once_1980), toplam_ara_1980_2000 = sum(ara_1980_2000), toplam_sonra_2000 = sum(sonra_2000))

    # Bar chart çizimi
    ggplot(vis2, aes(x = ilce_adi.x)) +
      geom_bar(aes(y = toplam_once_1980), stat = "identity", fill = "blue", position = "dodge") +
      geom_bar(aes(y = toplam_ara_1980_2000), stat = "identity", fill = "green", position = "dodge") +
      geom_bar(aes(y = toplam_sonra_2000), stat = "identity", fill = "red", position = "dodge") +
      labs(title = "Toplam Değerler - İlçelere Göre", x = "İlçe Adı", y = "Toplam Değer") +
      scale_fill_manual(values = c("Once 1980" = "blue", "Ara 1980-2000" = "green", "Sonra 2000" = "red"),
                        name = "Zaman Aralığı",
                        labels = c("Once 1980", "Ara 1980-2000", "Sonra 2000")) +
      theme_minimal() +
      theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))+
      guides(fill = guide_legend(title = "Zaman Aralığı"))
    ```

    **This plot shows us total building for each time zone vs district.**

    Note: blue --\> before 1980, green --\> between 1980 - 2000, red --\> after 2000.
